---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DFOsamplestats

<!-- badges: start -->
[![R-CMD-check](https://github.com/poissonconsulting/DFOsamplestats/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/poissonconsulting/DFOsamplestats/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/poissonconsulting/DFOsamplestats/branch/main/graph/badge.svg)](https://app.codecov.io/gh/poissonconsulting/DFOsamplestats?branch=main)
<!-- badges: end -->

The goal of DFOsamplestats is to ...

## Installation

`DFOsamplestats` can be installed from GitHub using the remotes package:

```r
install.packages("remotes")
remotes::install_github("poissonconsulting/DFOsamplestats")
```

## Usage

`DFOsamplestats` provides functions to calculate the power (at a significance level of 0.05) or sample size to achieve a power of 0.8 at a significance level of 0.05 using `stats::power.prop.test()`.
It is important to note that `n` is the total number of samples in the study (not the number of samples in each group).

```{r}
library(DFOsamplestats)
prop2_power(0.1, n = 10)
prop2_samplesize(0.1)
```
`DFOsamplestats` also provides functions to calculate the p-value and 95% confidence intervals based on the number of samples appearing in each group.
```{r}
prop2_pvalue(1, 9)
prop2_ci(1, 9)
```

`DFOsamplestats` also provides a function to simulate data.

```{r}
set.seed(42)
prop2_sim(0.1, 10)
set.seed(42)
prop2_sim(0.1, 10)
prop2_sim(0.1, 10)
```
which is called internally by a function which performs a full power analysis.
```{r}
prop2_power_analysis(0.1, 10)
```

It is worth noting that the power estimated by `prop2_power_analysis(0.1, 10)`
is slightly less than that estimated by `prop2_power(0.1, 10)`.
This is because `prop2_power()` uses a heuristic. 
To understand why `prop2_power()` is less reliable at small sample sizes
consider the fact that it is not possible to get a significant
result with just 5 samples.
```{r}
prop2_pvalue(0, 5)
prop2_power(0.1, 5)
prop2_power_analysis(0.1, 5)
```

This is because

## Contribution

Please report any [issues](https://github.com/poissonconsulting/DFOsamplestats/issues).

[Pull requests](https://github.com/poissonconsulting/DFOsamplestats/pulls) are always welcome.