---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DFOsamplestats

<!-- badges: start -->
[![R-CMD-check](https://github.com/poissonconsulting/DFOsamplestats/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/poissonconsulting/DFOsamplestats/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of DFOsamplestats is to provide functions to facilitate frequentist power analyses for return rates.

## Installation

`DFOsamplestats` can be installed from GitHub using the pak package:

```r
install.packages("pak")
pak::pak("poissonconsulting/DFOsamplestats")
```

## Usage

`DFOsamplestats` provides functions to calculate the power (at a significance level of 0.05) and sample size to achieve a power of 0.8 for the difference in the rate between two groups.

```{r}
library(DFOsamplestats)
rate2_power(p1 = 0.005, p2 = 0.0075, n = 31197)
rate2_samplesize(p1 = 0.005, p2 = 0.0075)
```

`DFOsamplestats` also provides functions to estimate the effect size and p-values 
based on the number of individuals returning in each group given the number tagged.

```{r}
rate_effect(r = c(156,  234), n = 31197)
```

And to simulate data

```{r}
set.seed(99)
rate_sim(p = c(0.005, 0.0075), n = 31197)
rate_sim(p = c(0.005, 0.0075), n = 31197)
```

Finally `DFOsamplestats` provides a function to perform a full power analysis.

```{r}
set.seed(100)
rate2_power_analysis(p1 = 0.005, p2 = 0.0075, n = 31197)
```

It is worth noting that `rate2_power()`, which is a wrapper on `stats::power.prop.test()`,
uses a heuristic to calculate the power.
As a result it is quicker but less reliable at small sample sizes than `rate2_power_analysis()` which performs a full power analysis (although calculation of the p-values still assumes that the likelihood profile is normally distributed).

To understand why `rate2_power()` is unreliable at small sample sizes consider the fact that it is not possible to get a significant result with just 5 samples yet it estimates the power to be 0.39 with 5 samples and a rate of 0.1!

```{r}
rate_effect(0, 5)
rate2_power(0.1, 5)
rate2_power_analysis(0.1, 5)
```

## Contribution

Please report any [issues](https://github.com/poissonconsulting/DFOsamplestats/issues).

[Pull requests](https://github.com/poissonconsulting/DFOsamplestats/pulls) are always welcome.
